NDNS multilevel LCA CW2CB2 or CW2CB3

### 2018-08-07 paper/Chapters/Chapter2.tex
want to figure out how to do model diagnosis

\subsection{Socio-demographic status, lifestyle, physical activity, anthropometric measurements and biochemical analysis}\vspace{-0.3cm}

@@

\vspace{-0.5cm}
\subsection{Non-Parametric approach}\vspace{-0.3cm}

@@

Multiple linear regression model (for WC), multinomial logistic regression model (for BMI categories), logistic regression models (for hypertension, and diabetes) were used to test for associations between individual level carbohydrate eating latent classes and WC, obesity, hypertension, and diabetes, separately.

Model fitting strategies

>>>>

\subsection{Socio-demographic status, lifestyle, physical activity, anthropometric measurements and biochemical analyses}\vspace{-0.3cm}

@@

\vspace{-0.5cm}
\subsubsection{Non-Parametric approach}\vspace{-0.3cm}

@@

Associations between individual level carbohydrate eating classes and the prevalence of hypertension, obesity as well as diabetes were explored in men and women separately. Point estimates of weighted means and proportions and 95\%CI of the characteristics were determined by applying either nurse visiting weights (for outcomes of hypertension, BMI, and WC) or blood sample weights (for outcomes of DM) accordingly. Similarly, the \textit{F} test (for continuous variables) and adjusted Pearson $\chi^2$ test (for categorical variables) were used to determine sex-specific differences by hypertension status, BMI categories, and DM diagnosis.

Logistic regression models (for hypertension, and diabetes), multinomial logistic regression model (for BMI categories), and linear regression model (for WC), were used to test for associations between individual level carbohydrate eating classes and hypertension, obesity, WC, and diabetes, separately.

For the multiple regression models, model fitting strategies are as follows:

\begin{enumerate}
	\item The crude association between the carbohydrate eating groups and the outcomes was first examined. 
	\item Potential confounders of the association between carbohydrate eater groups (exposure) and the outcomes were selected depending on the descriptive statisitics conducted above, ie. those are associated with both the exposure and the outcome and not on the causal pathway were selected as potential factors. 
	\item Confounding and/or interaction effect from each of the potential factors were checked one by one. 
	\item A preliminary model that includes all of the variables suggested to be confounders in the previous step was established. 
	\item The remaining variables were added to the preliminary model one by one to see if any of them may be a confounder in condition of the presence of the other covariates. 
	\item For logistic regression models, Goodness-of-fit (calibration) was assessed using the adapted \code{svylogitgof} command in Stata \parencite{archer2006goodness}. Other diagnostics (residual check,  Quadratic and cubic polynomial functions of continuous variables (such as age) were also tested to allow for nonlinear effects of these covariates on the binary outcomes.

	\item For linear regression model (WC), assumptions were checked visually by QQ plots of the residuals (normality), plotting the residuals against fitted values (constant variance). Outliers, leverage, and Cook's distance were checked visually to see if there may be 1) any extreme values of the dependent variable; 2) any points that have extreme values of the predictor variables; and 3) points which exert particular influence on the parameters of the fitted regression model. 
	\item Since likelihood ratio test is not appropriate under survey designed regression models, adjusted Wald test with $p < 0.05$ were used as criteria for variable inclusion in the final model. 
\end{enumerate}

Data manipulation and preparation \textbf{(Appendix \ref{AppendixA})} were done in R version 3.5.1 \parencite{R3.5.1}. All statistical analyses, except for MLCA models, were performed with "svyset" command as implemented in Stata software version 15.1 \parencite{stata15}. All \textit{P} values were two-sided.

### 2018-08-07 paper/Chapters/Chapter3.tex
want to figure out how to do model diagnosis

A series of traditional LCA of the responses to carbohydrate intake within 7 time slots of day was first examined. These initial analyses ignored the clustering of observation days within participants of the survey. \textbf{Table \ref{tab:mixmodels}} shows the latent class solutions for one to five classes (see rows under the Fixed effects model section). The BIC declines with the number of day level classes increases. However, the improvement of BIC dropped to less than 1000 from 3 classes to 4 classes solutions (658.9) and from 4 classes to 5 classses solutions (361.7). Entropy index indicates that the 4 classes model could explain about 51\% percent of the data, while \textit{p} values of Lo-Mendell-Rubun LRT suggest that the more classes we fit, the better model we will have until up to 6 classes (\textit{p} = 0.06 and is not shown in the table). However, from the parsimony point of view, we extend the model with random effects building on 2 classes, 3 classes and 4 classes solutions.

@@

In this 3 classes solution we have chosen in the day level, there were 39.5\%, 20.4\%, and 40.1\% classified into 3 latent groups. The overall counts and percentages for each responses within every time slot and the distributions of the solution are presented in \textbf{Table \ref{tab:daylevel}}. The trajectories illustrating the change of the probabilities of each response to carbohydrate eating during the hours of the day are shown separately by the three types of days in \textbf{Figure \ref{fig:level1}}.

@@

Class 1 days \textbf{(Figure \ref{fig:level1}-A)} were given the name of "high carbohydrate day" since in these days of survey, the probabilities of carbohydrate contributed higher or equal to 50\% of the energy consumed were always higher than that in the other two types of days. Specifically, these days were characterised with probabilities of over 0.6 in time slots between 6 am to 9 am, 9 am to 12 am, and also 2 pm to 5 pm, during which the time slots may be interpreted as breakfast, morning snack, and afternoon snack time periods for many participants. Moreover, even during late night time period, such as 8 pm to 10 pm, and 10 pm to 6 am time slots, the probabilities of having higher carbohydrate contained food were still as high as 0.412, and 0.246, respectively.

@@

Class 2 days \textbf{(Figure \ref{fig:level1}-B)} were named as "low carbohydrate day" because first of all, in these days the possibility of participants skipping breakfast was 0.45. And after 9 am, within class 2 days, the probability of having food with lower carbohydrate (contributed less than 50\% of total energy intake), was always higher than having higher carbohydrate contained food. In these days, participants also turned to have morning snacks (with only 0.079 possibility of Not eating any food and similar probabilities of having either high or low carbohydrate contained food). This phenomenon may also be interpreted as having a long and late breakfast in these mornings. The probability of Not eating any food was lowest for class 2 days during the midnights time slot (10 pm to 6 am), with about 0.590 compared with 0.680 and 0.751 during the class 1 and class 3 days.

@@

Class 3 days \textbf{(Figure \ref{fig:level1}-C)} were called "regular meals day" due to the following reasons: 1) participants recordings showed that in these days there was almost 0 possibility of not eating any food at lunch (0.019 between 12 noon and 2 pm) or dinner (0.034 between 5 pm and 8 pm); 2) the probabilities of not having morning snack (9 am to 12 am) and afternoon snack (2 pm to 5 pm) were also the highest among the three types of days (0.401 and 0.659). 3) during these days, participants may have some high carbohydrate contained food between 8 pm and 10 pm (0.308), but the probability of not eating any food during 10 pm to 6 am next morning was 0.751, the highest among the three types of days. \vspace{-0.5cm}

@@

As expected, consumption of total energy (7539.98 kJ), total carbohydrate (222.79 g), total sugar (98.63 g), starch (124.07 g), and non-milk extrinsic sugar (59.45 g) were highest among high carbohydrate days (all \textit{p} < 0.0001). On the other hand, the consumption of protein (73.29 g), total fat (67.24 g), and alcohol (13.80 g) were the highest in the so-called regular meals days. Moreover, in high carbohydrate days, participants turned to also consume the highest amount of fruit (107.40 g). There was no evidence of any difference for the consumption of yellow, red, or green vegetables across the three types of days (\textit{p} = 0.681).

@@

With three individual level latent classes \textbf{(Figure \ref{fig:level2})}, a low-carbohydrate eaters class, a moderate-carbohydrate eaters class, and a high-carbohydrate eaters class emerges. 43.1\% participants were identified as high-carbohydrate eaters, in these individuals, about 50\% of the days (2 out of 4 days) of their dietary diary were classified as having high carbohydrate days. Nearly 1 out of 4 days of their dietary diary were either "regular meals day" or "low carbohydrate day". 28.1\% participants fell into the low carbohydrate eaters class in the left hand side of \textbf{Figure \ref{fig:level2}}, their recordings of food intake showed that in more than 60\% of their days, they were having "regular meals" which was characterised as with highest amount of fat and alcohol consumptions as already described in \textbf{Table \ref{tab:day-level-features}}.

@@

The social-demographic characteristics of the UK adults according to their individual level latent class membership are shown in \textbf{Table \ref{tab:Level2tab1}}. Moderate carbohydrate eaters were relatively younger (\textit{p} < 0.001), and slightly less from England (\textit{p} = 0.007). Gender distribution across the three types of carbohydrate eaters were fairly even (\textit{p} = 0.119). Distribution of the carbohydrate eater types was changing with the year of survey. Low carbohydrate eaters represented 32.5\% of the population in the first year of survey, but later dropped to lower than 30\% (lowest in the third year, 22.6\%) until the most recent year. Proportion of high carbohydrate eaters increased from 41.2\% to the highest (50.6\%) in the second year of the survey, but then started to decline to 38.4\% in the 8th year of survey (\textit{p} = 0.015). There was no evidence of

@@

\caption{\label{tab:Level2tab1}Weighted means, percentages, and
95\% CIs of the social-demographic characteristics by latent class memberships in the UK
adults. (NDNS RP 2008/09-2015/16, sample size = 6155)} \centering
\fontsize{9}{11}\selectfont

\begin{tabular}[t]{lcccc}
	\hiderowcolors
	\toprule
	\textbf{Variables} & \textbf{\Centerstack{Low carbo-\\hydrate eaters\\(n = 1730)}} & \textbf{\Centerstack{Moderate carbo-\\hydrate eaters\\(n = 1772)}} & \textbf{\Centerstack{High carbo-\\hydrate eaters\\(n = 2653)}} & \textbf{\textit{P} value} \textsuperscript{*}\\
	\midrule
	\showrowcolors
	Total (\%) & 28.4 (26.9, 29.9) & 28.7 (27.1, 30.3) & 43.0 (41.3, 44.7) & \\
	Country (\%) &  &  &  & 0.007\\
	\hspace{1em}England & 84.5 (81.7, 86.9) & 82.0 (79.3, 84.5) & 84.7 (82.3, 86.8) & \\
	\hspace{1em}Northern Ireland & 2.1 (1.6, 2.8) & 4.2 (3.2, 5.6) & 2.2 (1.7, 3.0) & \\
	\hspace{1em}Scotland & 9.1 (7.0, 11.8) & 8.6 (6.7, 11.1) & 8.0 (6.3, 10.2) & \\
	\hspace{1em}Wales & 4.3 (3.3, 5.6) & 5.1 (4.0, 6.4) & 5.1 (4.0, 6.4) & \\
	Age (years) & 51.0 (49.9, 52.1) & 40.3 (39.1, 41.6) & 51.7 (50.7, 52.7) & < 0.001\\
	Sex (\%) &  &  &  & 0.119\\
	\hspace{1em}Men & 50.0 (46.9, 53.1) & 50.2 (47.0, 53.5) & 46.6 (44.0, 49.1) & \\
	\hspace{1em}Women & 50.0 (46.9, 53.1) & 49.8 (46.5, 53.0) & 53.4 (50.9, 56.0) & \\
	Survey years (\% in rows) &  &  &  & 0.015\\
	\hspace{1em}1 & 32.5 (28.4, 36.9) & 26.3 (21.9, 31.2) & 41.2 (36.6, 46.0) & \\
	\hspace{1em}2 & 26.8 (22.6, 31.3) & 22.6 (18.6, 27.3) & 50.6 (45.8, 55.4) & \\
	\hspace{1em}3 & 22.6 (18.8, 26.9) & 33.7 (28.6, 39.2) & 43.6 (38.7, 48.7) & \\
	\hspace{1em}4 & 27.9 (24.1, 32.2) & 27.6 (23.8, 31.8) & 44.4 (40.2, 48.7) & \\
	\hspace{1em}5 & 27.9 (24.2, 32.0) & 28.7 (24.4, 33.5) & 43.3 (38.2, 48.6) & \\
	\hspace{1em}6 & 28.0 (24.0, 32.4) & 31.5 (26.9, 36.6) & 40.5 (35.8, 45.3) & \\
	\hspace{1em}7 & 29.1 (25.2, 33.4) & 29.0 (24.5, 34.0) & 41.8 (37.1, 46.7) & \\
	\hspace{1em}8 & 31.1 (27.3, 35.3) & 30.5 (25.9, 35.5) & 38.4 (34.1, 42.8) & \\
	Paid employment\textsuperscript{\dag} (\%) &  &  &  & 0.907\\
	\hspace{1em}Yes & 40.3 (37.0, 43.6) & 40.8 (37.1, 44.5) & 39.8 (37.1, 42.6) & \\
	\hspace{1em}No & 59.7 (56.4, 63.0) & 59.2 (55.5, 62.9) & 60.2 (57.4, 62.9) & \\
	Live with partner\textsuperscript{\ddag} (\%) &  &  &  & < 0.001\\
	\hspace{1em}Yes & 56.9 (53.6, 60.1) & 38.4 (35.2, 41.8) & 61.3 (58.7, 63.7) & \\
	\hspace{1em}No & 43.1 (39.9, 46.4) & 61.6 (58.2, 64.8)) & 38.7 (36.3, 41,3) & \\
	Household Income, \textsterling/year & \Centerstack{36558.5\\(34800.2, 38316.8)} & \Centerstack{27180.8\\(25597.9, 28763.7)} & \Centerstack{32171.6\\(31024.9, 33318.2)} & < 0.001\\
	Ethnicity (\%) &  &  &  & \\
	\hspace{1em}White & 94.2 (92.4, 95.6) & 79.5 (76.4, 82.3) & 91.9 (90.1, 93.4) & < 0.001\\
	\hspace{1em}Non-White & 5.8 (4.4, 7.6) & 20.5 (17.7, 23.6) & 8.1 (6.6, 9.9) & \\
	Education (\%) &  &  &  & \\
	\hspace{1em}Degree or higher & 29.0 (26.1, 32.1) & 23.3 (20.5, 26.3) & 26.2 (24.1, 28.5) & 0.019\\
	\hspace{1em}Lower than degree & 71.0 (67.9, 73.9) & 76.7 (73.7, 79.5) & 73.8 (71.5, 75.9) & \\
	\bottomrule
	\multicolumn{5}{l}{{\scriptsize \textit{Note: }}}\\
	\multicolumn{5}{l}{{\scriptsize \textbf{Abbreviations}: CI, confidence intervals; NDNS RP, national dietary and nutrition survey rolling programme.}}\\
	\multicolumn{5}{l}{{\scriptsize Variables were weighted by individual weights.}}\\
	\multicolumn{5}{l}{{\scriptsize \textsuperscript{*} For continuous variables, the \textit{F} test was used to determine differences between latent classes with Bonferroni}}\\ 
	\multicolumn{5}{l}{{\scriptsize correction to account for multiple testing across $>$ 2 classes. For categorical variables, differences between}}\\
	\multicolumn{5}{l}{{\scriptsize  latent classes were assessed using the adjusted Pearson $\chi^2$ test for survey data.}} \\
	\multicolumn{5}{l}{{\scriptsize \textsuperscript{\dag} Paid employment was defined as being in paid employment during the last 4 weeks prior to the survey.} }\\
	\multicolumn{5}{l}{{\scriptsize \textsuperscript{\ddag} Live with partner was defined as either living with a married husband/wife or a legally recognised civil}} \\
	\multicolumn{5}{l}{{\scriptsize partnership.}}\\
\end{tabular}

@@

difference in employment status across three types of carbohydrate eaters. However, strong evidence suggested that high carbohydrate eaters had the highest proportion (61.3\%) of living with a partner (\textit{p} < 0.001); moderate carbohydrate eaters had the lowest average income (27180.8 \textsterling/year), highest proportion of non-white population (20.5\%), and lower education level (23.3\% with degree of higher education) compared with either low or high carbohydrate eaters.

Weighted means, percentages of anthropometric measurements, average of main nutrient intake, as well as biochemical characteristic profiles of latent carbohydrate eater groups are given in \textbf{Table \ref{tab:tab2}}. Low carbohydrate eaters were with higher BMI (27.8 kg/m\textsuperscript{2}) and larger WC (98.9/89.9 cm in men/women) compared with 27.2, 27.3 kg/m\textsuperscript{2}, and 95.9/88.7 (men/women), 98.1/87.2 (men/women) cm in moderate and high carbohydrate eaters. Moderate carbohydrate eaters had the highest prevalence of being a current smoker (27.8\%), shortest physical activity (geometric mean: 0.87 hours/day), and the lowest prevalence of hypertension (20.2\%).

Average total energy intake over the 4 days of dietary survey was the highest (7985.8 kJ) in the low carbohydrate eaters group. As expected, the mean total carbohydrate intake was 203.8 g, 218.3 g, and 233.4 g for low, moderate, and high carbohydrate eaters, respectively. Energy contribution from carbohydrate was close to 50\% in the high carbohydrate eaters, but was only 40.6\% in the low carbohydrate eaters. It is noteworthy that low carbohydrate eaters consumed the highest average amount of protein (79.9 g, 17.2\% of total energy), fat (74.7g, 35.4\% of total energy), and alcohol (20.8 g, 6.8\% of total energy).

From the results of blood tests, 6.9\% of low carbohydrate eaters were found to be diabetes (diagnosed by A1C > 6.5\%), while the percentages of diabetes in the moderate and high carbohydrate eaters were 3.5\%, and 4.1\% (\textit{p} < 0.011). There were some evidence (\textit{p} = 0.027) that fasting blood glucose level may be slightly higher in non-diabetes low carbohydrate eaters, while the geometric mean for A1C was probably lower in moderate carbohydrate eaters (4.72, 95\%CI: 5.39, 5.47). Cholesterol, HDL, and LDL were all lower in the moderate carbohydrate eaters, while no evidence of any difference of TG was found across three types of carbohydrate eaters.

@@

	\caption{\label{tab:tab2}Weighted means, percentages, and 95\% CIs of the anthropometric measurements, average main nutrients intake and biochemical characteristics by latent class memberships in the UK adults. \\(NDNS RP 2008/09-2015/16, sample size = 6155).}

@@

@@

@@

@@

>>>>

A series of traditional LCA of the responses to carbohydrate intake within 7 time slots of day was first examined. These initial analyses ignored the clustering of observation days within participants of the survey. \textbf{Table \ref{tab:mixmodels}} shows the latent class solutions for one to five classes (see rows under the Fixed effects model section). The BIC declines with the number of day level classes increases. However, the improvement of BIC dropped to less than 1000 from 3 classes to 4 classes solutions (658.9) and from 4 classes to 5 classses solutions (361.7). Entropy index indicates that the 4 classes model could explain about 51\% percent of the data, while \textit{p} values of Lo-Mendell-Rubun LRT suggest that the more classes we fit, the better model we will have until up to 6 classes (\textit{p} = 0.06 and is not shown in the table). From the parsimony point of view, we extended the model with random effects building on 2 classes, 3 classes and 4 classes solutions.

@@

In the 3$\times$3 random effect model solution we have chosen, there were 39.5\%, 20.4\%, and 40.1\% observations classified into 3 latent groups in the day level. The overall counts and percentages for each responses within every time slot and the distributions of the solution are presented in \textbf{Table \ref{tab:daylevel}}. The trajectories illustrating the change of the probabilities of each response to carbohydrate eating during the hours of the day are shown separately by the three types of days in \textbf{Figure \ref{fig:level1}}.

@@

Class 1 days \textbf{(Figure \ref{fig:level1}-A)} were given the name of "high carbohydrate day" since in these days of survey, the probabilities of carbohydrate contributed higher or equal to 50\% of the energy consumed were always higher than that in the other two types of days. Specifically, high carbohydrate days were characterised with probabilities of over 0.6 in time slots between 6 am to 9 am, 9 am to 12 am, and also 2 pm to 5 pm, during which the time slots may be interpreted as breakfast, morning snack, and afternoon snack time periods for many participants. Moreover, even during late night time period, such as 8 pm to 10 pm, and 10 pm to 6 am time slots, the probabilities of having higher carbohydrate contained food were still as high as 0.412, and 0.246, respectively.

@@

Class 2 days \textbf{(Figure \ref{fig:level1}-B)} were named as "low carbohydrate day" because first of all, in these days the possibility of participants skipping breakfast was 0.45. And after 9 am, within low carbohydrate days, the probability of having food contained lower carbohydrate (contributed less than 50\% of total energy intake), was always higher than having higher carbohydrate contained food. In these days, participants also turned to have morning snacks (with only 0.079 possibility of not eating any food and similar probabilities of having either high or low carbohydrate contained food). This phenomenon may also be interpreted as having a long and late breakfast (brunch) in these mornings. The probability of not eating any food was the lowest for low carbohydrate days during the midnight time slot (10 pm to 6 am), with probability of 0.590 compared with 0.680 and 0.751 in the class 1 and class 3 days, respectively.

@@

Class 3 days \textbf{(Figure \ref{fig:level1}-C)} were called "regular meals day" due to the following reasons: 1) participants' dietary recordings showed that in these days there was almost 0 possibility of not eating any food at lunch (0.019 between 12 noon and 2 pm) and dinner (0.034 between 5 pm and 8 pm); 2) the probabilities of not eating during morning snack time (9 am to 12 am) and afternoon snack time (2 pm to 5 pm) were also the highest among the three types of days (0.401 and 0.659). 3) during these days, participants may have some high carbohydrate contained food between 8 pm and 10 pm (0.308), but the probability of not eating any food during 10 pm to 6 am next morning was 0.751, the highest among the three types of days. \vspace{-0.5cm}

@@

As expected, consumption of total energy (7539.98 kJ), total carbohydrate (222.79 g), total sugar (98.63 g), starch (124.07 g), and non-milk extrinsic sugar (59.45 g) were highest among high carbohydrate days (all \textit{p} < 0.001). On the other hand, the consumption of protein (73.29 g), total fat (67.24 g), and alcohol (13.80 g) were the highest in the so-called regular meals days. Moreover, in high carbohydrate days, participants turned to consume the highest amount of fruit (107.40 g). There was no evidence of any difference for the consumption of yellow, red, or green vegetables across the three types of days (\textit{p} = 0.681).

@@

With three individual level latent classes \textbf{(Figure \ref{fig:level2})}, a low-carbohydrate eaters class, a moderate-carbohydrate eaters class, and a high-carbohydrate eaters class emerges. 43.1\% participants were identified as high-carbohydrate eaters, in these individuals, about 50\% of the days (2 out of 4 days) of their dietary diary could be classified as having high carbohydrate days. Nearly 1 out of 4 days of their dietary diary were either "regular meals day" or "low carbohydrate day". 28.1\% participants fell into the low carbohydrate eaters class in the left hand side of \textbf{Figure \ref{fig:level2}}, their recordings of food intake showed that in more than 60\% of their days, they were having "regular meals" which was characterised as with highest amount of fat and alcohol consumptions as already described in \textbf{Table \ref{tab:day-level-features}}. Moderate carbohydrate eaters have comparable proportions (42.0\% vs. 40.0\%) of having high carbohydrate days and regular meals day, 18.0\% of their dietary diary were identified as low carbohydrate days.

@@

The social-demographic characteristics of the UK adults according to their individual level latent class membership are shown in \textbf{Table \ref{tab:Level2tab1}}. Moderate carbohydrate eaters were relatively younger (\textit{p} < 0.001), and slightly less from England (\textit{p} = 0.007). Gender distribution across the three types of carbohydrate eaters was fairly even (\textit{p} = 0.119). Distribution of the carbohydrate eater types turned out to be changing with the year of survey. Low carbohydrate eaters represented 32.5\% of the population in the first year of survey, but later dropped to lower than 30\% (lowest in the third year, 22.6\%) until the most recent year. Proportion of high carbohydrate eaters increased from 41.2\% to the highest (50.6\%) in the second year of the survey, but then started to decline to 38.4\% in the 8th year of survey (\textit{p} = 0.015). There was no evidence of difference in employment status across three types of carbohydrate eaters. However, strong evidence suggested that high carbohydrate eaters had the highest proportion (61.3\%) of living with a partner (\textit{p} < 0.001); moderate carbohydrate eaters had the lowest average income (27180.8 \textsterling/year), highest proportion of non-white population (20.5\%), and lower education level (23.3\% with degree of higher education) compared with either low or high carbohydrate eaters.

Weighted means, percentages of anthropometric measurements, average of main nutrients intake, as well as biochemical characteristic profiles according to the latent carbohydrate eater groups are given in \textbf{Table \ref{tab:tab2}}. Low carbohydrate eaters had higher mean BMI (27.8 kg/m\textsuperscript{2}) and larger mean WC (98.9/89.9 cm in men/women) compared with 27.2, 27.3 kg/m\textsuperscript{2}, and 95.9/88.7 (men/women), 98.1/87.2 (men/women) cm in moderate and high carbohydrate eaters. Moderate carbohydrate eaters had the highest prevalence of being a current smoker (27.8\%), shortest time of daily physical activity (geometric mean: 0.87 hours/day), and the lowest prevalence of hypertension (20.2\%).

Average total energy intake over the 4 days of dietary survey was the highest (7985.8 kJ) in the low carbohydrate eaters group. As expected, the mean of total carbohydrate intake was 203.8 g, 218.3 g, and 233.4 g for low, moderate, and high carbohydrate eaters, respectively. Energy contribution from carbohydrate was close to 50\% in the high carbohydrate eaters, but was only 40.6\% in the low carbohydrate eaters. It is noteworthy that low carbohydrate eaters consumed the highest average amount of protein (79.9 g, 17.2\% of total energy), fat (74.7g, 35.4\% of total energy), and alcohol (20.8 g, 6.8\% of total energy).

From the results of blood tests, 6.9\% of low carbohydrate eaters were found to be diabetic (diagnosed by A1C > 6.5\%), while the percentages of diabetes in the moderate and high carbohydrate eaters were 3.5\%, and 4.1\% (\textit{p} < 0.011), respectively. Although there was some evidence (\textit{p} = 0.027) that fasting blood glucose level may be slightly higher in non-diabetic low carbohydrate eaters, the geometric mean for A1C was probably lower in moderate carbohydrate eaters (4.72, 95\%CI: 5.39, 5.47). Cholesterol, HDL, and LDL were all lower in the moderate carbohydrate eaters, while no evidence of any difference of TG was found across three types of carbohydrate eaters.

@@

\caption{\label{tab:Level2tab1}Weighted means, percentages, and
95\% CIs of the social-demographic characteristics by carbohydrate eating latent class memberships in the UK
adults. \\ (NDNS RP 2008/09-2015/16, sample size = 6155)} \centering
\fontsize{9}{11}\selectfont

\begin{tabular}[t]{lcccc}
	\hiderowcolors
	\toprule
	\textbf{Variables} & \textbf{\Centerstack{Low carbo-\\hydrate eaters\\(n = 1730)}} & \textbf{\Centerstack{Moderate carbo-\\hydrate eaters\\(n = 1772)}} & \textbf{\Centerstack{High carbo-\\hydrate eaters\\(n = 2653)}} & \textbf{\textit{P} value} \textsuperscript{*}\\
	\midrule
	\showrowcolors
	Total (\%) & 28.4 (26.9, 29.9) & 28.7 (27.1, 30.3) & 43.0 (41.3, 44.7) & \\
	Country (\%) &  &  &  & 0.007\\
	\hspace{1em}England & 84.5 (81.7, 86.9) & 82.0 (79.3, 84.5) & 84.7 (82.3, 86.8) & \\
	\hspace{1em}Northern Ireland & 2.1 (1.6, 2.8) & 4.2 (3.2, 5.6) & 2.2 (1.7, 3.0) & \\
	\hspace{1em}Scotland & 9.1 (7.0, 11.8) & 8.6 (6.7, 11.1) & 8.0 (6.3, 10.2) & \\
	\hspace{1em}Wales & 4.3 (3.3, 5.6) & 5.1 (4.0, 6.4) & 5.1 (4.0, 6.4) & \\
	Age (years) & 51.0 (49.9, 52.1) & 40.3 (39.1, 41.6) & 51.7 (50.7, 52.7) & < 0.001\\
	Sex (\%) &  &  &  & 0.119\\
	\hspace{1em}Men & 50.0 (46.9, 53.1) & 50.2 (47.0, 53.5) & 46.6 (44.0, 49.1) & \\
	\hspace{1em}Women & 50.0 (46.9, 53.1) & 49.8 (46.5, 53.0) & 53.4 (50.9, 56.0) & \\
	Survey years (\% in rows) &  &  &  & 0.015\\
	\hspace{1em}1 & 32.5 (28.4, 36.9) & 26.3 (21.9, 31.2) & 41.2 (36.6, 46.0) & \\
	\hspace{1em}2 & 26.8 (22.6, 31.3) & 22.6 (18.6, 27.3) & 50.6 (45.8, 55.4) & \\
	\hspace{1em}3 & 22.6 (18.8, 26.9) & 33.7 (28.6, 39.2) & 43.6 (38.7, 48.7) & \\
	\hspace{1em}4 & 27.9 (24.1, 32.2) & 27.6 (23.8, 31.8) & 44.4 (40.2, 48.7) & \\
	\hspace{1em}5 & 27.9 (24.2, 32.0) & 28.7 (24.4, 33.5) & 43.3 (38.2, 48.6) & \\
	\hspace{1em}6 & 28.0 (24.0, 32.4) & 31.5 (26.9, 36.6) & 40.5 (35.8, 45.3) & \\
	\hspace{1em}7 & 29.1 (25.2, 33.4) & 29.0 (24.5, 34.0) & 41.8 (37.1, 46.7) & \\
	\hspace{1em}8 & 31.1 (27.3, 35.3) & 30.5 (25.9, 35.5) & 38.4 (34.1, 42.8) & \\
	Paid employment\textsuperscript{\dag} (\%) &  &  &  & 0.907\\
	\hspace{1em}Yes & 40.3 (37.0, 43.6) & 40.8 (37.1, 44.5) & 39.8 (37.1, 42.6) & \\
	\hspace{1em}No & 59.7 (56.4, 63.0) & 59.2 (55.5, 62.9) & 60.2 (57.4, 62.9) & \\
	Live with partner\textsuperscript{\ddag} (\%) &  &  &  & < 0.001\\
	\hspace{1em}Yes & 56.9 (53.6, 60.1) & 38.4 (35.2, 41.8) & 61.3 (58.7, 63.7) & \\
	\hspace{1em}No & 43.1 (39.9, 46.4) & 61.6 (58.2, 64.8)) & 38.7 (36.3, 41,3) & \\
	Household income, \textsterling/year & \Centerstack{36558.5\\(34800.2, 38316.8)} & \Centerstack{27180.8\\(25597.9, 28763.7)} & \Centerstack{32171.6\\(31024.9, 33318.2)} & < 0.001\\
	Ethnicity (\%) &  &  &  & \\
	\hspace{1em}White & 94.2 (92.4, 95.6) & 79.5 (76.4, 82.3) & 91.9 (90.1, 93.4) & < 0.001\\
	\hspace{1em}Non-White & 5.8 (4.4, 7.6) & 20.5 (17.7, 23.6) & 8.1 (6.6, 9.9) & \\
	Education (\%) &  &  &  & \\
	\hspace{1em}Degree or higher & 29.0 (26.1, 32.1) & 23.3 (20.5, 26.3) & 26.2 (24.1, 28.5) & 0.019\\
	\hspace{1em}Lower than degree & 71.0 (67.9, 73.9) & 76.7 (73.7, 79.5) & 73.8 (71.5, 75.9) & \\
	\bottomrule
	\multicolumn{5}{l}{{\scriptsize \textit{Note: }}}\\
	\multicolumn{5}{l}{{\scriptsize \textbf{Abbreviations}: CI, confidence intervals; NDNS RP, national dietary and nutrition survey rolling programme.}}\\
	\multicolumn{5}{l}{{\scriptsize Variables were weighted by individual weights.}}\\
	\multicolumn{5}{l}{{\scriptsize \textsuperscript{*} For continuous variables, the \textit{F} test was used to determine differences between latent classes with Bonferroni}}\\ 
	\multicolumn{5}{l}{{\scriptsize correction to account for multiple testing across $>$ 2 classes. For categorical variables, differences between}}\\
	\multicolumn{5}{l}{{\scriptsize  latent classes were assessed using the adjusted Pearson $\chi^2$ test for survey data.}} \\
	\multicolumn{5}{l}{{\scriptsize \textsuperscript{\dag} Paid employment was defined as being in paid employment during the last 4 weeks prior to the survey.} }\\
	\multicolumn{5}{l}{{\scriptsize \textsuperscript{\ddag} Live with partner was defined as either living with a married husband/wife or a legally recognised civil}} \\
	\multicolumn{5}{l}{{\scriptsize partnership.}}\\
\end{tabular}

@@

@@

	\caption{\label{tab:tab2}Weighted means, percentages, and 95\% CIs of the anthropometric measurements, average main nutrients intake and biochemical characteristics by carbohydrate eating latent class memberships in the UK adults. (NDNS RP 2008/09-2015/16, sample size = 6155)}

@@

		Household income,\textsterling/year & 34006.5 (31972.9, 36040.1) & 32280.5 (29875.6, 34685.4) & 0.284 &  32741.5 (31009.9, 34473.1) & 27862.0 (25557.0, 30167.0)  &  < 0.001  \\
		Live with partner\textsuperscript{\ddag}, Yes, (\%) & 56.1 (51.8, 61.4) & 66.6 (61.3, 71.5) & 0.002   & 48.7 (45.1, 52.3) & 58.9 (53.6, 63.9)  & 0.002    \\

@@

		DM\textsuperscript{\S} (\%) & 3.7 (2.4, 5.7) & 12.6 (8.9, 17.5) & < 0.001 & 1.8 (1.0, 3.3) & 7.9 (5.1, 11.9) & < 0.001 \\ 

@@

		\multicolumn{7}{l}{\textsuperscript{\ddag} Live with partner was defined as either living with a married husband/wife or a legally recognised civil partnership.}\\

@@

		\multicolumn{7}{l}{\textsuperscript{\S}  DM was defined by A1C $>$ 6.5\%.}
